---
title: "Homework 2"
output: github_document
author: Sushupta Vijapur
---
```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

Read in the Mr.Trashwheel dataset

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```
The columns 15-17 were added bc of the unnamed notes column

Read precipitation data 2017 and 2018 and clean.

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine precipitation datasets
month.names are character vectors built into R
   mutate(month = month.name[month])
   But this is ugly and we don't like it
Below is a better way.
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2017, precip_2018)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore Maryland. As trash enters the inner harbor, the trashwheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additionally, data sheets include month precipitation data. 

# Problem 2

## Part a
Load NYC transit data frame and clean it. 

```{r}
nyc_transit = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")   %>%
  janitor::clean_names() %>% 
  select(line:route11, entry, vending, entrance_type, ada) %>% 
  mutate(
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"))
```

## Part b

Descriptive understanding of the data frame.

```{r}
nrow(distinct(nyc_transit, line, station_name))

filter(nyc_transit, ada == "TRUE")

filter(nyc_transit, vending == "NO")
```

This dataset contains information related to each entrance and exit for each subway station in NYC. There are `r nrow(nyc_transit)` in the observations in the data set and 19 variables. In order to clean the data, I used the "clean_names" function in the janitor package to change all variable naming mechanisms to snake case. Furthermore, I only selected in relevant variables into the data frame (line, station name, station latitude / longitude, routes served, entry, vending, entrance type and ADA compliance). Furthermore, I recoded the entry variable from Yes/No to a logical as True/False. 

While running some descriptives on the dataset, I found that there was 465 distinct stations (defined by both the statio name and the line). Additionally, I found that there was 468 observations which are ADA compliant

# Problem 3

Load pols-month csv file.

```{r}
month_df = 
  tibble(
    month = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12"),
    month_name = month.name
  )

pols_month = 
  read_csv("./data/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"))

pols_month_df =
  left_join(pols_month, month_df) %>% 
  relocate(month_name) %>%
  mutate(
    president = recode(
      prez_dem, `1` = "dem", `0` = "gop")) %>% 
   select(-month, -prez_dem, -prez_gop, -day)

  

```


