---
title: "Homework 2"
output: github_document
author: Sushupta Vijapur
---
```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

Read in the Mr.Trashwheel dataset and tidy data.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
    sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```
The columns 15-17 were added because of the unnamed notes column at the end of the dataset.

Read precipitation data 2017 and 2018 and clean.

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine precipitation datasets
month.names are character vectors built into R
   mutate(month = month.name[month])
   But there is a better way below.
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2017, precip_2018)

left_join(precip_df, month_df, by = "month")
```


Now, join precipitation dataframs with the trashwheel dataframes.
```{r}
prob1_df = 
  left_join(trashwheel_df, precip_df, by = "year")
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore Maryland. As trash enters the inner harbor, the trashwheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additionally, data sheets include month precipitation data. The total preciptiation in 2018 was `r sum(pull(precip_2018, total))` inches. Also, the median number of sports balls in a dumpster in 2017 was `r median(pull(prob1_df, sports_balls))`. 

# Problem 2

## Part a

Load NYC transit data frame and clean it. 

```{r}
nyc_transit = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")   %>%
  janitor::clean_names() %>% 
  select(line:route11, entry, vending, entrance_type, ada) %>% 
  mutate(
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"))
```

This dataset contains information related to each entrance and exit for each subway station in NYC. There are `r nrow(nyc_transit)` in the observations in the data set and 19 variables. In order to clean the data, I used the "clean_names" function in the janitor package to change all variable naming mechanisms to snake case. Furthermore, I only selected in relevant variables into the data frame (line, station name, station latitude / longitude, routes served, entry, vending, entrance type and ADA compliance). Furthermore, I recoded the entry variable from Yes/No to a logical as True/False. The data are not tidy yet since the route data (numbers and names) are listed in a wide format and have many NA values.

Run code to figure out descriptives in data.

```{r}
no_vending_df = nyc_transit %>% 
  filter(vending == "NO")
prop_entrance = round(nrow(filter(no_vending_df, entry == TRUE)) / nrow(no_vending_df),3)


```

While running some descriptives on the dataset, (1) I found that there were `r count(distinct(nyc_transit,line, station_name))` distinct stations (defined by both the station name and the line). (2) Additionally,  there were `r count(no_vending_df, ada = TRUE)` observations which were ADA compliant. (3) Lastly, among the entrance/exits that don't allow vending, 37.7% of them allow entry.

## Part b

Cleaning and pivoting route number and names.

```{r}
nyc_transit_tidy = 
nyc_transit %>% 
  mutate(
    route8 = as.character(route8),
    route9 = as.character(route9),
    route10 = as.character(route10),
    route11 = as.character(route11)) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    names_prefix = "route",
    values_to = "route_name") %>% 
  drop_na("route_name")

```

The data are tidy after I reformatted the route data to list route number and route name in a long format instead of a wide format.There were `r nyc_transit_tidy %>% filter(route_name == "A") %>% distinct(line, station_name, route_name == "A") %>% count()` distinct stations that served the A train. Furthermore, of the stations that served the A train, `r nyc_transit_tidy %>% filter(route_name == "A" & ada =="TRUE") %>% distinct(line, station_name, route_name == "A") %>% count()` were ADA compliant.


# Problem 3

Load and clean/tidy pols-month csv file.

```{r}
month_pols = 
  tibble(
    month = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12"),
    month_name = month.name
  )

pols_month = 
  read_csv("./data/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"))

pols_month_tidy =
  left_join(pols_month, month_pols) %>% 
  relocate(month_name) %>%
  mutate(
    president = ifelse(prez_dem == 1, "dem", "gop")) %>% 
   select(-month, -prez_dem, -prez_gop, -day)
```

Load and clean/tidy the snp csv file.

```{r}
month_snp = 
  tibble(
    month = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"),
    month_name = month.name
  )

snp = 
  read_csv("./data/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"))

snp_tidy = 
  left_join(snp, month_snp, by = "month") %>% 
  select(-month, -day) %>% 
  relocate(month_name)
```

Load and clean/tidy unemployment csv.

```{r}
month_unemploy = 
  tibble(
    month = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"),
    month_name = month.name
  )

unemploy = 
  read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment"
  ) 

unemploy_tidy = 
  left_join(unemploy, month_unemploy, by = "month") %>% 
  select(-month) %>% 
  relocate(month_name) %>% 
  mutate(year = as.character(year))
```

Join the datasets together. 

```{r}
prob3_df = 
  left_join(pols_month_tidy, snp_tidy, by = c('month_name' = 'month_name', 'year' = 'year')) %>% 
  left_join(unemploy_tidy, by = c('month_name' = 'month_name', 'year' = 'year'))
```

The final merged "prob3_df" includes the "pols_month_tidy", "snp_tidy" and "unemploy_tidy" dataframes. The "pols_month_tidy" dataframe contains 822 observations and 11 variables. I recoded the date variable to be separate month and year columns. Also, I created a president column and used "prez_gop" and "prez_dem" to denote whether they were democrates or republicans. Furthermore, the dataset includes variables for the number of democratic/republican governors, senators and representatives. The "snp_tidy" dataframe has 787 observations. Similar to in the pols_month_tidy dataframe, I separated the month and year variables into 2 separate columns. Lastly, the "unemploy_tidy" dataframe has 816 observations (after we pivoted the original file to long from wide). There is also a month and year column along with a percent column indicating the percentage of unemployment in each month of a certain year. 

Once we merged these three dataframes, the final dataset contains `r nrow (prob3_df)` and `r ncol(prob3_df)`. This final dataframe contains all the same columns as in the 3 merged dataframes. The year range in this dataframe is `r range(as.numeric(pull(prob3_df, year)))`.

