---
title: "Homework2"
output: github_document
---
```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read in the Mr.Trashwheel dataset

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```
The columns 15-17 were added bc of the unnamed notes column

Read precipitation data 2017 and 2018

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine precipitation datasets
month.names are character vectors built into R
   mutate(month = month.name[month])
   But this is ugly and we don't like it
Below is a better way
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2017, precip_2018)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore Maryland. As trash enters the inner harbor, the trashwheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additionally, data sheets include month precipitation data. 