---
title: "Homework 2"
output: github_document
author: Sushupta Vijapur
---
```{r setup}
library(tidyverse)
library(readxl)
```

# Problem 1

Read in the Mr.Trashwheel dataset and tidy data.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-7-2020-1.xlsx",
    sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```
The columns 15-17 were added because of the unnamed notes column at the end of the dataset.

Read precipitation data 2017 and 2018 and clean.

```{r}
precip_2018 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 = 
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine precipitation datasets
month.names are character vectors built into R
   mutate(month = month.name[month])
   But there is a better way below.
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2017, precip_2018)

left_join(precip_df, month_df, by = "month")
```

This dataset contains information from the Mr.Trashwheel trash collector in Baltimore Maryland. As trash enters the inner harbor, the trashwheel collects that trash and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific types of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additionally, data sheets include month precipitation data. 

# Problem 2

## Part a
Load NYC transit data frame and clean it. 

```{r}
nyc_transit = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")   %>%
  janitor::clean_names() %>% 
  select(line:route11, entry, vending, entrance_type, ada) %>% 
  mutate(
    entry = recode(entry, "YES" = "TRUE", "NO" = "FALSE"))
```

## Part b

Descriptive understanding of the data frame. (1) Finding out how many distinct stations there are and (2) how many stations are ADA compliant. (3) Lastly, finding the proportion of station entrances / exits without vending that allow entrance.

```{r}



nyc_transit %>% 
  count(entry, vending) 
```

This dataset contains information related to each entrance and exit for each subway station in NYC. There are `r nrow(nyc_transit)` in the observations in the data set and 19 variables. In order to clean the data, I used the "clean_names" function in the janitor package to change all variable naming mechanisms to snake case. Furthermore, I only selected in relevant variables into the data frame (line, station name, station latitude / longitude, routes served, entry, vending, entrance type and ADA compliance). Furthermore, I recoded the entry variable from Yes/No to a logical as True/False. The data here are tidy after making these adjustments to the formatting of the dataframe.

While running some descriptives on the dataset, (1) I found that there were `r count(distinct(nyc_transit,line, station_name))` distinct stations (defined by both the station name and the line). (2) Additionally,  there were 468 observations which were ADA compliant. (3) Lastly, among the entrance/exits that don't allow vending, 37.7% of them allow entry.

# Problem 3

Load and clean/tidy pols-month csv file.

```{r}
month_pols = 
  tibble(
    month = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12"),
    month_name = month.name
  )

pols_month = 
  read_csv("./data/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"))

pols_month_tidy =
  left_join(pols_month, month_pols) %>% 
  relocate(month_name) %>%
  mutate(
    president = recode(
      prez_dem, `1` = "dem", `0` = "gop")) %>% 
   select(-month, -prez_dem, -prez_gop, -day)
```

Load and clean/tidy the snp csv file.

```{r}
month_snp = 
  tibble(
    month = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"),
    month_name = month.name
  )

snp = 
  read_csv("./data/snp.csv") %>% 
  separate(date, into = c("month", "day", "year"))

snp_tidy = 
  left_join(snp, month_snp, by = "month") %>% 
  select(-month, -day) %>% 
  relocate(month_name)
```

Load and clean/tidy unemployment csv.

```{r}
month_unemploy = 
  tibble(
    month = c("jan", "feb", "mar", "apr", "may", "jun", "jul", "aug", "sep", "oct", "nov", "dec"),
    month_name = month.name
  )

unemploy = 
  read_csv("./data/unemployment.csv") %>% 
  janitor::clean_names() %>%
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "percent"
  ) 

unemploy_tidy = 
  left_join(unemploy, month_unemploy, by = "month") %>% 
  select(-month) %>% 
  relocate(month_name) %>% 
  mutate(year = as.character(year))
```

Join the datasets together. 

```{r}
prob3_df = 
  left_join(pols_month_tidy, snp_tidy, by = c('month_name' = 'month_name', 'year' = 'year')) %>% 
  left_join(unemploy_tidy, by = c('month_name' = 'month_name', 'year' = 'year'))
```

The final merged "prob3_df" includes the "pols_month_tidy", "snp_tidy" and "unemploy_tidy" dataframes. The "pols_month_tidy" dataframe contains 822 observations. I recoded the date variable to be separate month and year columns. Also, I created a president column and used "prez_gop" and "prez_dem" to denote whether they were democrates or republicans. Furthermore, the dataset includes variables for the number of democratic/republican governors, senators and representatives. The "snp_tidy" dataframe gass 787 observations. Similar to in the pols_month_tidy dataframe, I separated the month and year variables into 2 separate columns. Lastly, the "unemploy_tidy" dataframe has 816 observations (after we pivoted the original file to long from wide). There is also a month and year column along with a percent column indicating the percentage of unemployment in each month of a certain year. 

Once we merged these three dataframes, the final dataset contains `r nrow (prob3_df)` and `r ncol(prob3_df)`. This final dataframe contains all the same columns as in the 3 merged dataframes. The year range in this dataframe is `r range(as.numeric(pull(prob3_df, year)))`.

